---
title: "CellsOfInterest"
output: html_document
date: '2022-05-30'
---
Load packages
```{r, include = FALSE}
# install.packages("Seurat") - V4.1
library(Seurat)
# install.packages("ggplot2") - V3.3.6
library(ggplot2)
# install.packages("tidyverse") - V1.3.2
library(tidyverse)

memory.limit(size=40000)
# to always have the same parameters 
set.seed(2)

# Specific for HPC to make plots
options(bitmapType = "cairo")
```

Read SeuratObject: 
```{r}
setwd("/kyukon/scratch/gent/gvo000/gvo00097/vsc43612/2022_SingleCell_Stéphane/RDSObjects")
ada_stefNoALK <- readRDS("ada_and_stefNoALK_dim20_res06.rds")
```

Show UMAP - Adameyko + Stéphane
```{r}
DimPlot(ada_stefNoALK, reduction="umap", group.by = "seurat_clusters", label = TRUE) + 
  ggtitle("Adameyko + Stéphane noALK")
```

Show UMAP - Adameyko and Stéphane seperated
```{r}
DimPlot(ada_stefNoALK, reduction="umap", split.by = "nez", label = TRUE)

# Keep only data from Stéphane 
stéphane_only <- subset(ada_stefNoALK, subset = nez == "sc")
saveRDS("stéphane_only_SLB.rds")
```

Focus only on the cells of interest (Schwann cell precursors, Bridging cells, Chromaffin like cells, sympathoblasts, cycling sympathoblasts)
```{r}
cells_of_interest <- subset(stéphane_only, idents = c("9", "11", "12", "18", "22"))
DimPlot(cells_of_interest, reduction="umap", group.by = "seurat_clusters", label = TRUE)
ggsave(file = "../DataAnalysis_SLB/CellsOfInterest/OnlyCells_Of_Interest.pdf", device = "pdf")

# Calculate the percentage cells per cluster
percentages_per_cluster <- as.data.frame(cells_of_interest@meta.data) %>% group_by(seurat_clusters) %>% summarise(ncells_cluster=n()) %>% mutate(prop_cluster=ncells_cluster/sum(ncells_cluster))


percentages_per_cluster

percentages_per_cluster %>%
  ggplot(aes(x=seurat_clusters, y=prop_cluster, fill = seurat_clusters)) + 
  geom_bar(stat="identity") + 
  ggtitle("Cells of interest")
ggsave(file = "../DataAnalysis_SLB/CellsOfInterest/Barplot_OnlyCells_Of_Interest.pdf", device = "pdf")
```

Redo the normalization and clustering steps only on your cells of interest

Based on: DefaultAssay(cells_of_interest)<- "integrated"
```{r}
DefaultAssay(cells_of_interest)<- "integrated"
cells_of_interest <- FindVariableFeatures(object = cells_of_interest, selection.method = "vst", nfeatures = 2000)

### Plot variable features with and without labels
top10 <- head(x = VariableFeatures(object = cells_of_interest), 10)
plot1 <- VariableFeaturePlot(object = cells_of_interest)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)

print(plot2)
```

Scaling Data

```{r, message=FALSE, warning=FALSE,tidy=T}
cells_of_interest <- ScaleData(object = cells_of_interest, features = rownames(cells_of_interest))

##### Check per group #####
cells_of_interest[['nUMI_scaled']]<-colSums(as.matrix(cells_of_interest[['integrated']]@scale.data))
cells_of_interest[['nGene_scaled']]<-apply(as.matrix(cells_of_interest[['integrated']]@scale.data),2,function(x){sum(x>0)})

VlnPlot(cells_of_interest, features=c("nUMI_scaled","nGene_scaled"))
```

PCA analysis after scaling

```{r, message=FALSE, warning=FALSE,tidy=T}
cells_of_interest <- RunPCA(object = cells_of_interest, features = VariableFeatures(cells_of_interest), npcs = 60, ndims.print = 1:5, nfeatures.print = 10)
#seuratObj[['pca']]@cell.embeddings[1:5,1:5]

```

Significant PCs

```{r, message=FALSE, warning=FALSE,tidy=T}

### Run Jackstraw and create plot. This can take a couple of minutes to calculate
cells_of_interest <- JackStraw(object = cells_of_interest, num.replicate = 100, dims = 60)
cells_of_interest <- ScoreJackStraw(object = cells_of_interest, dims = 1:60)

JackStrawPlot(object = cells_of_interest, dims = 1:60)
ElbowPlot(object = cells_of_interest, ndims = 60)
```

Clustering cells 
```{r, message=F, warning=F, tidy=T}
#save old clusters
cells_of_interest@meta.data$seurat_clusters_old <- cells_of_interest@meta.data$seurat_clusters

# Play with these variables until you have a good tSNE/clustering
dimsToTry <- 20
resToUse <- 0.3

##### Find clusters
cells_of_interest <- FindNeighbors(object = cells_of_interest, dims = 1:dimsToTry)
cells_of_interest <- FindClusters(object = cells_of_interest, resolution = resToUse)

##### Create tSNE plot
cells_of_interest <- RunTSNE(object = cells_of_interest, dims = 1:dimsToTry,check_duplicates = FALSE)
tsnePlot <- DimPlot(cells_of_interest, reduction = "tsne", label = T, label.size = 8, pt.size = 1)
tsnePlot

##### Create UMAP plot
cells_of_interest<- RunUMAP(cells_of_interest, dims = 1:dimsToTry, n_neighbors = 30)
umapPlot <- DimPlot(cells_of_interest, reduction = "umap", label = T, label.size = 8)

umapPlot
# New UMAP after normalization and clustering but coloured using old colors 
DimPlot(cells_of_interest, reduction = "umap", label = T, label.size = 8, group.by = "seurat_clusters_old") + 
  ggtitle("Cells of Interest after redoing normalization and clustering")
ggsave(file = "../DataAnalysis_SLB/CellsOfInterest/UMAP_Cells_Of_Interest_afterRedoingNormClust_OldColors_Integrated.pdf", device = "pdf")

# UMAP with new colors en clusters
DimPlot(cells_of_interest, reduction = "umap", label = T, label.size = 8, group.by = "seurat_clusters" ) + 
  ggtitle("Cells of Interest after redoing normalization and clustering")
ggsave(file = "../DataAnalysis_SLB/CellsOfInterest/UMAP_Cells_Of_Interest_afterRedoingNormClust_NewColors_Integrated.pdf", device = "pdf")
```

FeaturePlot 
```{r}
# Schwann Cell Precursor - SOX10
FeaturePlot(cells_of_interest, features = c("SOX10"))
ggsave(file = "../DataAnalysis_SLB/CellsOfInterest/FeaturePlot_SOX10_NewClustering_Integrated.pdf", device = "pdf")

# Bridging cells - ASCL1
FeaturePlot(cells_of_interest, features = c("ASCL1"))
ggsave(file = "../DataAnalysis_SLB/CellsOfInterest/FeaturePlot_ASCL1_NewClustering_Integrated.pdf", device = "pdf")

# Chromaffin like-progenitor cells - DBH
FeaturePlot(cells_of_interest, features = c("DBH"))
ggsave(file = "../DataAnalysis_SLB/CellsOfInterest/FeaturePlot_DBH_NewClustering_Integrated.pdf", device = "pdf")

# Sympatho-adrenergic precursors - STMN2
FeaturePlot(cells_of_interest, features = c("STMN2"))
ggsave(file = "../DataAnalysis_SLB/CellsOfInterest/FeaturePlot_STMN2_NewClustering_Integrated.pdf", device = "pdf")

# Cycling Sympatho-adrenergic precursors - TOP2A
FeaturePlot(cells_of_interest, features = c("TOP2A"))
ggsave(file = "../DataAnalysis_SLB/CellsOfInterest/FeaturePlot_TOP2A_NewClustering_Integrated.pdf", device = "pdf")

```

```{r}
# Save data 
saveRDS(cells_of_interest,"CellsOfInterest_SLB.rds")
```

Check differential expressed genes between clusters. 

```{r}
Idents(cells_of_interest) <- "seurat_clusters"

# Cluster 0 vs all 
cluster0_vs_all <- FindMarkers(cells_of_interest, ident.1 = 0)
head(cluster0_vs_all, n = 5)

write.table(cluster0_vs_all, "../DataAnalysis_SLB/CellsOfInterest/DifferentialExpressedGenes_Cluster0_vs_AllClusters.txt")

# Cluster 1 vs all 
cluster1_vs_all <- FindMarkers(cells_of_interest, ident.1 = 1)
head(cluster1_vs_all, n = 5)

write.table(cluster1_vs_all, "../DataAnalysis_SLB/CellsOfInterest/DifferentialExpressedGenes_Cluster1_vs_AllClusters.txt")

# Cluster 2 vs all 
cluster2_vs_all <- FindMarkers(cells_of_interest, ident.1 = 2)
head(cluster2_vs_all, n = 5)

write.table(cluster2_vs_all, "../DataAnalysis_SLB/CellsOfInterest/DifferentialExpressedGenes_Cluster2_vs_AllClusters.txt")

# Cluster 3 vs all 
cluster3_vs_all <- FindMarkers(cells_of_interest, ident.1 = 3)
head(cluster3_vs_all, n = 5)

write.table(cluster3_vs_all, "../DataAnalysis_SLB/CellsOfInterest/DifferentialExpressedGenes_Cluster3_vs_AllClusters.txt")

# Cluster 4 vs all 
cluster4_vs_all <- FindMarkers(cells_of_interest, ident.1 = 4)
head(cluster4_vs_all, n = 5)

write.table(cluster4_vs_all, "../DataAnalysis_SLB/CellsOfInterest/DifferentialExpressedGenes_Cluster4_vs_AllClusters.txt")

# Cluster 5 vs all 
cluster5_vs_all <- FindMarkers(cells_of_interest, ident.1 = 5)
head(cluster5_vs_all, n = 5)

write.table(cluster5_vs_all, "../DataAnalysis_SLB/CellsOfInterest/DifferentialExpressedGenes_Cluster5_vs_AllClusters.txt")

# Cluster 6 vs all 
cluster6_vs_all <- FindMarkers(cells_of_interest, ident.1 = 6)
head(cluster6_vs_all, n = 5)

write.table(cluster6_vs_all, "../DataAnalysis_SLB/CellsOfInterest/DifferentialExpressedGenes_Cluster6_vs_AllClusters.txt")

# Cluster 7 vs all 
cluster7_vs_all <- FindMarkers(cells_of_interest, ident.1 = 7)
head(cluster7_vs_all, n = 5)

write.table(cluster7_vs_all, "../DataAnalysis_SLB/CellsOfInterest/DifferentialExpressedGenes_Cluster7_vs_AllClusters.txt")
```

Barplots for the cells of interest 
```{r}
# Barplots for all cells
percentages_per_cluster <- as.data.frame(stéphane_only@meta.data) %>% group_by(seurat_clusters) %>% summarise(ncells_cluster=n()) %>% mutate(prop_cluster=ncells_cluster/sum(ncells_cluster))


percentages_per_cluster

percentages_per_cluster %>%
  ggplot(aes(x=seurat_clusters, y=prop_cluster, fill = seurat_clusters)) + 
  geom_bar(stat="identity") + 
  ggtitle("Only Stéphane")

# Barplot for cells of interest 
cells <- percentages_per_cluster[which(percentages_per_cluster$seurat_clusters == "12" | percentages_per_cluster$seurat_clusters == "22" |
                                   percentages_per_cluster$seurat_clusters == "11"| percentages_per_cluster$seurat_clusters == "9" |
                                     percentages_per_cluster$seurat_clusters == "18"),]

# Specific order for paper
cells %>%
  ggplot(aes(x=seurat_clusters, y=prop_cluster, fill = seurat_clusters)) + 
  geom_bar(stat="identity") + 
  ggtitle("Cells of interest with %cells of big UMAP") + 
  scale_x_discrete(limits = c("12", "22", "11", "9", "18")) 
ggsave(file = "../DataAnalysis_SLB/Barplots/Barplot_Cells_of_Interest_wittperfBigUMAP.pdf", device = "pdf")

# Barplot for extended cells of interest  
cells_ext <- percentages_per_cluster[which(percentages_per_cluster$seurat_clusters == "12" | percentages_per_cluster$seurat_clusters == "22" |
                                   percentages_per_cluster$seurat_clusters == "11"| percentages_per_cluster$seurat_clusters == "9" |
                                     percentages_per_cluster$seurat_clusters == "18" | percentages_per_cluster$seurat_clusters == "5" |
                                     percentages_per_cluster$seurat_clusters == "3"| percentages_per_cluster$seurat_clusters == "29"| 
                                     percentages_per_cluster$seurat_clusters == "13"| percentages_per_cluster$seurat_clusters == "2" |
                                     percentages_per_cluster$seurat_clusters == "4"),]

cells_ext

# Specific order for paper
cells_ext %>%
  ggplot(aes(x=seurat_clusters, y=prop_cluster, fill = seurat_clusters)) + 
  geom_bar(stat="identity") + 
  ggtitle("Cells of interest with %cells of big UMAP") + 
  scale_x_discrete(limits = c("12", "22", "11", "9", "18", "5", "3", "29", "13", "2", "4")) 
ggsave(file = "../DataAnalysis_SLB/Barplots/Barplot_Cells_of_Interest_EXT_wittperfBigUMAP.pdf", device = "pdf")



```

Dag 26 vs day 32
```{r}
# Now the difference between day 26 and day 32 for EACH specific cluster 
stéphane_only@meta.data$Days<- gsub(".*-D","",stéphane_only@meta.data$hash.ID)
DimPlot(stéphane_only, reduction="umap", group.by = "Days")

# Calculate % of cells per timepoint 
percentages_per_day <- as.data.frame(stéphane_only@meta.data) %>% 
  group_by(seurat_clusters, Days) %>% 
  summarise(ncells_cluster=n())
  
sum_all_celles = sum(percentages_per_day$ncells_cluster)  

percentages_per_day$prop_cluster <- percentages_per_day$ncells_cluster/sum_all_celles

# Correct! Sum of the different % of cells per timepoint is the same as above 
percentages_per_day

cells_days <- percentages_per_day[which(percentages_per_day$seurat_clusters == "12" | percentages_per_day$seurat_clusters == "22" |
                                   percentages_per_day$seurat_clusters == "11"| percentages_per_day$seurat_clusters == "9" |
                                     percentages_per_day$seurat_clusters == "18"),]

# For day 26! 
cells_day26 <- cells_days[which(cells_days$Days == "26"),]

cells_day26 %>%
  ggplot(aes(x=seurat_clusters, y=prop_cluster, fill = seurat_clusters)) + 
  geom_bar(stat="identity") + 
  ggtitle("Cells of interest with %cells of big UMAP for day 26") +
  scale_x_discrete(limits = c("12", "22", "11", "9", "18")) + 
  ylim(0,0.015)
ggsave(file = "../DataAnalysis_SLB/Barplots/Barplot_Cells_of_Interest_wittperfBigUMAP_day26.pdf", device = "pdf")

# For day 32! 
cells_day32 <- cells_days[which(cells_days$Days == "32"),]

cells_day32 %>%
  ggplot(aes(x=seurat_clusters, y=prop_cluster, fill = seurat_clusters)) + 
  geom_bar(stat="identity") + 
  ggtitle("Cells of interest with %cells of big UMAP for day 32") +
  scale_x_discrete(limits = c("12", "22", "11", "9", "18")) + 
  ylim(0,0.015)
ggsave(file = "../DataAnalysis_SLB/Barplots/Barplot_Cells_of_Interest_wittperfBigUMAP_day32.pdf", device = "pdf")

# Barplot for extended cells of interest 
cells_days_ext <- percentages_per_day[which(percentages_per_day$seurat_clusters == "12" | percentages_per_day$seurat_clusters == "22" |
                                              percentages_per_day$seurat_clusters == "11"| percentages_per_day$seurat_clusters == "9" |
                                              percentages_per_day$seurat_clusters == "18" | percentages_per_day$seurat_clusters == "5" |
                                              percentages_per_day$seurat_clusters == "3"| percentages_per_day$seurat_clusters == "29"|
                                              percentages_per_day$seurat_clusters == "13"| percentages_per_day$seurat_clusters == "13" |
                                              percentages_per_day$seurat_clusters == "2" |percentages_per_day$seurat_clusters == "4"),]

cells_days_ext


# For day 26! 
cells_day26_ext <- cells_days_ext[which(cells_days_ext$Days == "26"),]

cells_day26_ext %>%
  ggplot(aes(x=seurat_clusters, y=prop_cluster, fill = seurat_clusters)) + 
  geom_bar(stat="identity") + 
  ggtitle("Cells of interest with %cells of big UMAP for day 26") +
  scale_x_discrete(limits = c("12", "22", "11", "9", "18", "5", "3", "29", "13", "2", "4")) + 
  ylim(0,0.015)
ggsave(file = "../DataAnalysis_SLB/Barplots/Barplot_Cells_of_Interest_EXT_wittperfBigUMAP_day26.pdf", device = "pdf")

# For day 32! 
cells_day32_ext <- cells_days_ext[which(cells_days_ext$Days == "32"),]

cells_day32_ext %>%
  ggplot(aes(x=seurat_clusters, y=prop_cluster, fill = seurat_clusters)) + 
  geom_bar(stat="identity") + 
  ggtitle("Cells of interest with %cells of big UMAP for day 32") +
  scale_x_discrete(limits = c("12", "22", "11", "9", "18", "5", "3", "29", "13", "2", "4")) + 
  ylim(0,0.015)
ggsave(file = "../DataAnalysis_SLB/Barplots/Barplot_Cells_of_Interest_EXT_wittperfBigUMAP_day32.pdf", device = "pdf")

